Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire \\[1cm] \huge Supplementary Material

### 2016-10-18 manuscript/supplementary.Rnw
Updates to manuscript

\documentclass[9pt,twocolumn,twoside, lineno]{gcb}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment. 
\pdfminorversion=4
\templatetype{gcbresearcharticle} % Choose template

% Change label of supplementary figures
\renewcommand{\thefigure}{S\arabic{figure}}

\title{Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b,c]{James S Camac}
\author[d]{Richard J Williams} 
\author[c]{Carl-Henrik Wahren}
\author[e]{Ary A Hoffmann}
\author[b]{Peter A Vesk}

\affil[a]{Department of Biological Sciences, Macquarie University, Sydney, NSW 2109, Australia}
\affil[b]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
\affil[c]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
\affil[d]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
\affil[e]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}

@@

\begin{document}
\suppmaketitle

\section*{SI Materials and Methods}
\subsection*{Study site}
Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals \citep{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.

\subsection*{Open Top Chamber Experiment}
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.

@@

We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient numbers of seedlings from species occurring in both open and closed heathland. We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all fire-killed (the dominant shrub post-fire strategy in the Australian mainland Alps; \citet{Walsh:2004vk}); common in the Australian Alps; and under climatic warming, are likely candidates to invading non-shrubby plant communities such as alpine grasslands and herbfields. Resprouter shrub species were not examined in this study because we were unable to locate sufficient numbers of seedlings; and are unlikely to be the primary contributors to shrub expansion in these landscapes. For each species, we collected seedlings with a range of heights (\Sexpr{range_init_size(complete_otc_seedling_data)[1]} to \Sexpr{range_init_size(complete_otc_seedling_data)[2]} cm) using a forestry planter (12.5 cm deep by 5.5 cm wide). The size of the planter was selected to minimize root damage across the range of seedling sizes and to allow for the extraction of a constant volume (378 cm$^3$) of intact soil and roots. The planter also minimized soil disturbance during seedling extraction and replanting.

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea} seedlings were transplanted into occuring post-fire inter-tussock gaps of varying size surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. \ref{figs8}). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

@@

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig. \ref{figs5}-\ref{figs7}). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsubsection*{Data inputs for growth, mortality and recruitment models}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations \citep{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

\subsection*{Seedling survey}
We measured seedling density and maximum heights in open heathland across the Bogong High Plains. The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (TWI; a measure of plant available water), adult density, fire, and fire severity influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland. A secondary aim was to verify growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.

@@

\subsubsection*{Data inputs into density and maximum height models}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations \citep{Gelman:2007te}.

\subsection*{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset \citep{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \citep{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

\subsubsection*{Modelling shrub seedling growth rates}
We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:

@@

\subsubsection*{Modelling shrub seedling mortality rates}
We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size. Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$. We modelled these binary observations through time as random realisations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:

@@

\subsubsection*{Modelling Grevillea recruitment rates}
We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:

@@

\subsubsection*{Modelling censored Poa distances}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e. $\geq$  26 cm).  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.

@@

\subsubsection*{Modelling rates of inter-tussock gap infilling by Poa}
As well as modelling how OTCs affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps. To examine this we first modelled the median observed gap radius, $y_{i,t}$, for inter-tussock gap $i$ in census $t$, as a random realisation from a lognormal distribution:

@@

\subsubsection*{Modelling drivers of seedling density}
We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:

@@

\subsubsection*{Modelling drivers of seedling maximum heights}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation \citep[equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C;][]{Brown:1989vx} influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:

@@

\bibliography{refs.bib}

\begin{figure*}
\centering
\includegraphics{figures/figS1.pdf}
\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI). (A) Centered and standardized model coefficients (on the log scale); and effects of (B) burning, (C) fire severity, (D) altitude and (E) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs1}
\end{figure*}
\clearpage

\begin{figure*}
\centering
\includegraphics{figures/figS2.pdf}
\caption{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
\label{figs2}
\end{figure*}

\clearpage
\begin{figure*}
\centering
\includegraphics{figures/figS3.pdf}
\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs. Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively. Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.} 
\label{figs3}
\end{figure*}
\clearpage

\begin{figure*}
\centering
\includegraphics{figures/figS4.pdf}
\caption{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires. Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. (A \& B) Centered and standardized model coefficients (on the log scale); and effects of (C \& D) fire severity, (E \& F) altitude and (G \& H) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs4}
\end{figure*}
\clearpage

\begin{figure*}
\centering
\includegraphics{figures/figS5.pdf}
\caption{Mean daily microclimate difference (OTC $-$ Control). (A) Ambient temperature, (B) soil temperature, (C) volumetric water content and (D) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs5}
\end{figure*}
\clearpage

\begin{figure*}
\centering
\includegraphics{figures/figS6.pdf}
\caption{Mean minimum daily microclimate difference (OTC $-$ Control). (A) Ambient temperature, (B) soil temperature, (C) volumetric water content and (D) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs6}
\end{figure*}
\clearpage

\begin{figure*}
\centering
\includegraphics{figures/figS7.pdf}
\caption{Mean maximum daily microclimate difference (OTC $-$ Control). Ambient temperature (A), soil temperature (B), volumetric water content (C) and relative humidity over experimental period (D). Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
\label{figs7}
\end{figure*}
\clearpage

\begin{figure*}
\centering
\includegraphics{figures/figS8.pdf}
\caption{Diagram outlining positions of transplanted seedling in an OTC plot. The inner white square represents the burnt bare ground patch used to assess heights of four dominant alpine shrub species. The grey area contains \textit{Grevillea australis} seedlings randomly transplanted into burnt inter-tussock gaps. The area beyond the grey section was subject to OTC panel sheltering (i.e. it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.} 
\label{figs8}
\end{figure*}
\clearpage

>>>>

\documentclass[9pt,onecolumn,twoside, lineno]{gcb}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment. 
\pdfminorversion=4
\templatetype{gcbresearcharticle} % Choose template 
\usepackage{float}

% Change label of supplementary figures
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\bibsection}{\section{\refname}}
\title{Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire \\[1cm]
\huge Supplementary Material}

% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
\author[a,b]{James S Camac}
\author[c]{Richard J Williams} 
\author[b]{Carl-Henrik Wahren}
\author[d]{Ary A Hoffmann}
\author[a]{Peter A Vesk}

\affil[a]{The Centre of Excellence for Environmental Decisions, School of BioSciences, The University of Melbourne, Melbourne, Victoria 3010, Australia}
\affil[b]{Research Centre for Applied Alpine Ecology, La Trobe University, Melbourne, Victoria 3086, Australia}
\affil[c]{CSIRO Land and Water, Tropical Ecosystems Research Centre, PMB 44 Winnellie, Northern Territory 0821, Australia}
\affil[d]{Bio21 Institute, School of BioSciences, The University of Melbourne, Victoria 3010, Australia}

@@

\begin{document}
\suppmaketitle
\tableofcontents
\listoffigures

\section{Study site}
Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals \citep{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.

\section{Open Top Chamber Experiment}
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.

@@

We collected seedlings from a site burnt by a late 2006 fire because it allowed us to collect sufficient numbers of seedlings from species occurring in both open and closed heathland. We focused on \textit{Asterolasia trymalioides} (Rutaceae; a prostrate shrub), \textit{Grevillea australis} (Proteaceae; a tall shrub), \textit{Phebalium squamulosum} (Rutaceae; a tall shrub) and \textit{Prostanthera cuneata} (Lamiaceae; a tall shrub) because they are all fire-killed \citep[the dominant shrub post-fire strategy in the Australian mainland Alps;][]{Walsh:2004vk}; common in the Australian Alps; and under climatic warming, are likely candidates to invading non-shrubby plant communities such as alpine grasslands and herbfields. Resprouter shrub species were not examined in this study because we were unable to locate sufficient numbers of seedlings; and are unlikely to be the primary contributors to shrub expansion in these landscapes. For each species, we collected seedlings with a range of heights (\Sexpr{range_init_size(complete_otc_seedling_data)[1]} to \Sexpr{range_init_size(complete_otc_seedling_data)[2]} cm) using a forestry planter (12.5 cm deep by 5.5 cm wide). The size of the planter was selected to minimize root damage across the range of seedling sizes and to allow for the extraction of a constant volume (378 cm$^3$) of intact soil and roots. The planter also minimized soil disturbance during seedling extraction and replanting.

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea} seedlings were transplanted into occuring post-fire inter-tussock gaps of varying size surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. \ref{figs1}). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

@@

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig. \ref{figs6}-\ref{figs8}). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsection{Data inputs for growth, mortality and recruitment models}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations \citep{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

\section{Seedling survey}
We measured seedling density and maximum heights in open heathland across the Bogong High Plains. The primary aim of this study was to evaluate how variables such as altitude, Topographic Wetness Index (TWI; a measure of plant available water), adult density, fire, and fire severity influence the density of \textit{Asterolasia} and \textit{Grevillea} seedlings in open heathland. A secondary aim was to verify growth rates observed in the OTC experiment by assessing maximum seedling heights of \textit{Asterolasia} and \textit{Grevillea} along altitude, soil moisture and fire severity gradients in areas burnt by the 2003 wildfires.

@@

\subsection{Data inputs into density and maximum height models}
In all seedling survey models, $Burnt_j$ is a binary (0 = unburnt, 1 = burnt) site level covariate, which only affects seedling counts at unburnt sites. By contrast, fire severity ($Severity_j$), was a continuous site level covariate that only affected seedling counts and maximum heights at burnt sites. Altitude ($Alt_j$), Topographic Wetness Index ($TWI_j$) and mature \textit{Grevillea} count ($ADen_k$) were also continuous covariates that affected all sites in both count and maximum height models. Collinearity among covariates was checked and all correlations were \textit{r} \textless 0.16. The data from each continuous covariate was then centered on their mean and standardized by two standard deviations \citep{Gelman:2007te}.

\section{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset \citep{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \citep{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

\subsection{Modelling shrub seedling growth rates}
We modelled the observed height of individual $i$ at May census $t$ as a random realisation from a normal distribution truncated at zero:

@@

\subsection{Modelling shrub seedling mortality rates}
We were also interested in whether the instantaneous mortality rates of seedlings varied amongst shrub species, OTC treatment and inter-tussock gap size. Our observations, $y_{i,t}$, were binary (0 = alive or 1 = dead) for each individual $i$, in each census, $t$. We modelled these binary observations through time as random realisations from a Bernoulli distribution based on the model-estimated cumulative hazard rate for each individual at each census:

@@

\subsection{Modelling Grevillea recruitment rates}
We examined whether warming influenced natural recruitment rates of \textit{Grevillea} by modelling the number of recruits per 1 m$^2$ plot $k$ in each census, $t$ as a random realisation from a negative binomial distribution:

@@

\subsection{Modelling censored Poa distances}
Approximately \Sexpr{prop_censored(cleaned_poa_distances)}\% of the $Poa$ distance data used to calculate the median inter-tussock gap radius, $Gap_{i,t}$, were recorded as censored observations, (i.e. $\geq$  26 cm).  We estimated these censored values by first estimating the distribution of distances in observed data, coupled with estimating covariate effects, and then using these distributions to estimate censored observations.

@@

\subsection{Modelling rates of inter-tussock gap infilling by Poa}
As well as modelling how OTCs affected shrub seedling growth and mortality rates, we were also interested in examining how warming influenced rates of grass infilling of inter-tussock gaps. To examine this we first modelled the median observed gap radius, $y_{i,t}$, for inter-tussock gap $i$ in census $t$, as a random realisation from a lognormal distribution:

@@

\subsection{Modelling drivers of seedling density}
We examined drivers of seedling density by modelling the number of seedlings per square metre plot $i$ as a random realisation from a Poisson distribution:

@@

\subsection{Modelling drivers of seedling maximum heights}
In order to verify growth rate findings in the OTC experiment, we assessed how a range of 190 m in elevation \citep[equivalent to a temperature change of \Sexpr{round(0.77 * (190/100),1)}\textdegree C;][]{Brown:1989vx} influenced \textit{Grevillea} and \textit{Asterolasia} seedling heights in areas burnt by the 2003 fires. We also assessed whether TWI and fire severity influenced these heights. Here, our observations were of maximum seedling heights for both species when at least one individual was present in a plot. We used maximum heights because these were likely to represent individuals that established immediately post-fire, and would have had 9 years of post-fire growth. We modelled these heights, $maxht_i$, as a random realisations from a log-normal distribution:

@@

\bibliography{refs.bib}
\pagebreak

\section{Supplementary Figures}

\begin{figure}[h]
\centering
\includegraphics{figures/figS1.pdf}
\caption{Diagram outlining positions of transplanted seedling in an OTC plot. The inner white square represents the burnt bare ground patch used to assess heights of four dominant alpine shrub species. The grey area contains \textit{Grevillea australis} seedlings randomly transplanted into burnt inter-tussock gaps. The area beyond the grey section was subject to OTC panel sheltering (i.e. it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.} 
\label{figs1}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS2.pdf}
\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI). (a) Centered and standardized model coefficients (on the log scale); and effects of (b) burning, (c) fire severity, (d) altitude and (e) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs2}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS3.pdf}
\caption{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
\label{figs3}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS4.pdf}
\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs. Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively. Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.} 
\label{figs4}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS5.pdf}
\caption{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires. Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. (a \& b) Centered and standardized model coefficients (on the log scale); and effects of (c \& d) fire severity, (e \& f) altitude and (g \& h) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs5}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS6.pdf}
\caption{Mean daily microclimate difference (OTC $-$ Control). (a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs6}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS7.pdf}
\caption{Mean minimum daily microclimate difference (OTC $-$ Control). (a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs7}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS8.pdf}
\caption{Mean maximum daily microclimate difference (OTC $-$ Control). Ambient temperature (a), soil temperature (b), volumetric water content (c) and relative humidity over experimental period (d). Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
\label{figs8}
\end{figure}
